---
title: 'Assignment 04: Tidy data and joins'
author: "Stanley Nam"
date: "October 8, 2019"
output: 
    html_document:
      keep_md: true
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
library(gapminder)
library(DT)
```

# Exercise 1: Univariate data reshaping

## Univariate Option 1

First, I am going to make a tibble with one row per year, and columns for life expectancy for France and Italy. I have chosen them as they are Romance-language-speaking countries with considerable consumption of red wine. Previous research has found out consumption of red wine reduces the rate of heart attack and thus it is expected to increase life expectancy. And then, I am going to use the new data shape to scatterplot life expectancy for one country against that of another. Finally, I am going to re-lengthen the data.

### 1. A tibble that has first column for year and two columns for the two countries.

```{r task 1.1}
FI_life <- gapminder %>%
  filter(country == "France" | country == "Italy") %>% 
  select(year, country, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)

FI_life
```

### 2. A scatterplot out of the tibble

```{r task 1.2}
FI_life %>%
  ggplot(aes(x = France, y = Italy)) +
  geom_point() +
  labs(title = "A Scatterplot of Life Expectancy between France and Italy",
       x = "France",
       y = "Italy") + 
  theme_bw()
```

### 3. Re-lengthening the data

```{r task 1.3}
FI_life %>% 
  pivot_longer(cols = c(France, Italy),
               names_to = "country",
               values_to = "lifeExp") %>% 
  arrange(country) %>% 
  datatable()
```


# Exercise 2: Multivariate data reshaping

## Multivariate Option 1

In this question, I am going to make a tibble with one row per year, and columns for life expectancy *and* GDP per capita for Oceanian countries (i.e., Australia and New Zealand).

After generating the tibble as above, I am going to re-lengthen the data.

### 1. A tibble with one column for year and other columns for the life expectancy and GDP per capita of the two countries

```{r task 2.1}
oceanianData <-
  gapminder %>% 
  filter(continent == "Oceania") %>% 
  select(year, country, lifeExp, gdpPercap) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep = "_",
              values_from = c(lifeExp, gdpPercap))

datatable(oceanianData)
```

### 2. Re-lengthening

```{r task 2.2}
oceanianData %>% 
  pivot_longer(cols = -year,
               names_to = c(".value", "country"),
               names_sep = "_") %>% 
  datatable()
```


# Exercise 3: Table Joins

```{r task 3.0, include=FALSE}
# Reading in the wedding guestlist (guest tibble) and email addresses(email tibble).
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## 3.1 

For each guest in the guestlist (*guest* tibble), add a column for email address, which can be found in the *email* tibble.

```{r task 3.1}
email <- email %>% 
  separate_rows(guest, sep = ', ') %>% # tibble email has different names concatenated in a row so separate them
  rename(name = guest)  # As in the guest tibble the names are under a column named 'name,' I'm using the same for the email tibble

guest %>% 
  left_join(email, by = "name")
```

## 3.2

Who do we have emails for, yet are not on the guestlist?

```{r task 3.2}
email %>%
  anti_join(guest, by = "name")
```

## 3.3

Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r task 3.3}
email %>%
  full_join(guest, by = "name") %>%
  select(-party) %>% # For some reason, we have another column called 'party.' Basically, persons with the same email share a value in this column. We don't need this so remove it.
  datatable()
```
