---
title: "HW02: Gapminder and dplyr"
author: "Stanley Nam"
date: "September 24, 2019"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages, warning=FALSE, echo=FALSE, message=F}
library(tidyverse)
library(gapminder)
```

# Exercise 1: Basic dplyr

## 1.1

Use filter() to subset the gapminder data to three countries of your choice in the 1970’s.

```{r 1.1}
subData <- gapminder %>%
  filter(country == "Philippines" | country == "Korea, Dem. Rep." | country == "Korea, Rep.") %>%
  filter(year >= 1970   & year <= 1979)

subData
```

## 1.2

Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1.

```{r 1.2}
subData %>% 
  select(country, gdpPercap)
```

## 1.3

Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that’s the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful.

```{r 1.3}
gapminder %>%
# the code below sorts gapminder by country and then year, so that data from a country are presented in increasing year
  arrange(country, year) %>%
  
# in the code below, new column 'lifeExpIncrease' tracks the increase in life expectancy by year.
  mutate(lifeExpIncrease = lifeExp - lag(lifeExp)) %>%
  
# in the code below, lifeExpIncrease below zero means a decrease in lifeExp; 'year - lag(year) > 0' is for sanity check. i.e., year - lag(year) is only above zero if we comparing lifeExp of a same country, as we sorted the data by country and then year.
  filter(lifeExpIncrease < 0, year - lag(year) > 0)
```

## 1.4

Filter gapminder so that it shows the max GDP per capita experienced by each country. Hint: you might find the max() function useful here.

```{r 1.4}
gapminder %>%
  group_by(country) %>%
  filter(gdpPercap == max(gdpPercap))
```

## 1.5

Produce a scatterplot of Canada’s life expectancy vs. GDP per capita using ggplot2, without defining a new variable. That is, after filtering the gapminder data set, pipe it directly into the ggplot() function. Ensure GDP per capita is on a log scale.

```{r 1.5}
gapminder %>%
  filter(country == "Canada") %>%
  ggplot(aes(gdpPercap, lifeExp)) +
  geom_point() +
  scale_x_log10('GDP per capita') +
  ylab("Life Expectancy") +
  theme_classic()
```

# Exercise 2: Explore individual variables with dplyr

Pick one categorical variable and one quantitative variable to explore. Answer the following questions in whichever way you think is appropriate, using dplyr:

> I choose *continent/country* as a categorical variable, and *pop(ulation)* as a quantitative variable.

## What are possible values (or range, whichever is appropriate) of each variable?

Possible values for *continent* are Africa, Americas, Asia, Europe and Oceania. 

```{r 2.1.1}
gapminder %>%
  select(continent) %>%
  summary()
```

Range of *population* is from 60011 to 1318683096.

```{r 2.1.2}
gapminder %>%
  select(pop) %>%
  range()
```

## What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.

Let's see the difference of the typical population (mean) of a country by continent.

```{r 2.1.3}
gapminder %>%
  group_by(continent) %>%
  mutate(contMean = mean(pop) / 1000000) %>%
  ggplot(aes(continent, contMean)) +
  scale_y_log10('Mean population (m)') +
  geom_point()
```

Now, compare this to the total population of each continent. By doing so, we can investigate whether a continent is overly broken down with country borders (whether there is too many countries in a continent).

```{r 2.1.4}
gapminder %>%
  group_by(continent) %>%
  mutate(contSum = sum(pop) / 1000000) %>%
  ggplot(aes(continent, contSum)) +
  scale_y_log10('Total population (m)') +
  geom_point()
```

We can conclude that Africa is overly broken down with borders.

# Exercise 3: Explore various plot types

Make two plots that have some value to them. That is, plots that someone might actually consider making for an analysis. Just don’t make the same plots we made in class – feel free to use a data set from the datasets R package if you wish.

## Correlation between increased life expectancy and population growth

```{r 3.1}
gapminder %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(lifeExpInc = tail(lifeExp, 1) - head(lifeExp, 1), popGrowth = (tail(pop, 1) - head(pop, 1)) / 1000000) %>%
  ggplot(aes(lifeExpInc, popGrowth)) +
  scale_y_log10('Population growth (m)') +
  xlab("Increase in life expectancy") +
  geom_point()
```

## How and when did South Korean economy (i.e., GDP per capita) overtake North Korea?

```{r 3.2}
gapminder %>%
  filter(country == "Korea, Dem. Rep." | country=="Korea, Rep.") %>%
  filter(year < 1985) %>%
  ggplot() +
  geom_line(aes(year, gdpPercap, group = country, color = country)) +
  xlab("Year") +
  ylab("GDP per capita") +
  theme(legend.position = "right")
```
