---
title: "HW05: Factor and figure management"
author: "Stanley Nam"
date: "October 17, 2019"
output: 
    html_document:
      keep_md: true
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
library(gapminder)
library(DT)
library(here)
```


# Exercise 1: Explain the value of the here::here package

For the cause of replicability, my data analysis should be disseminated with the intermediate process. So, in addition to the results and discussion section, I may provide the raw data and the R codes that I used to manipulate/analyze such data. However, not everyone has the same environment as I do: different OS's, different directory hierarchies (which makes things difficult to use absolute paths) and so one. Therefore, I should return the environment to the Tabula Rasa and fix it there before actually running the process. Previous practices, something that I actually used in a [previous project](), incorporate the use of the **rm()** and **setwd()** function. However, they simply mask the problem without providing a real solution! 

More problematic is some untold assumptions that were made in the process from my side. For example, previously loaded packages may have overwritten functions I intended to use! As such packages are not attached without an explicit statement, researchers trying to replicate my analysis would get different results. Therefore, it would be a better practice, not just for disseminating the data and analysis process but as a general habit in writing R codes, to start any code with the  **here** package.

# Exercise 2: Factor management

I chose to work on the *gapminder* dataset, and I am going to analyze *continent* (a factor).

DO SOMETHING HERE!!!!!!!!!!!!

# Exercise 3: File I/O
Experiment with at least one of:

- write_csv()/read_csv() (and/or TSV friends),
- saveRDS()/readRDS(),
- dput()/dget().


# Exercise 4: Visualization design

Create a side-by-side plot and juxtapose your first attempt

# Exercise 5: Writing figures to file

Use ggsave()
