---
title: "HW05: Factor and figure management"
author: "Stanley Nam"
date: "October 17, 2019"
output: 
    html_document:
      keep_md: true
      toc: true
      toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
library(gapminder)
library(DT)
library(here)
```


# Exercise 1: Explain the value of the here::here package

For the cause of replicability, my data analysis should be disseminated with the intermediate process. So, in addition to the results and discussion section, I may provide the raw data and the R codes that I used to manipulate/analyze such data. However, not everyone has the same environment as I do: different OS's, different directory hierarchies (which makes things difficult to use absolute paths) and so one. Therefore, I should return the environment to the Tabula Rasa and fix it there before actually running the process. Previous practices, something that I actually used in a [previous project](), incorporate the use of the **rm()** and **setwd()** function. However, they simply mask the problem without providing a real solution! 

More problematic is some untold assumptions that were made in the process from my side. For example, previously loaded packages may have overwritten functions I intended to use! As such packages are not attached without an explicit statement, researchers trying to replicate my analysis would get different results. Therefore, it would be a better practice, not just for disseminating the data and analysis process but as a general habit in writing R codes, to start any code with the  **here** package.

# Exercise 2: Factor management

I chose to work on the *gapminder* dataset. It has two factor variables *continent* and *country*. Between the two, my focus is on *continent*. Specifically, the datapoints associated with Oceania will be dropped.

Before we dive in the actual analysis, let's make sure that these two variables are actually factors.

```{r 2.0.1}
gapminder
```

It is explicitly shown that *country* and *continent* are actually <fctr> (factor) variables, and *continent* has five levels: "Africa", "Americas", "Asia", "Europe" and "Oceania" as shown below

```{r 2.0.2}
levels(gapminder$continent)
```

## 2.1 Drop factor / levels

The reason I chose to drop "Oceania" is simple: it has the fewest number of data points.

```{r 2.1}
no_oceania <- gapminder %>% 
  filter(continent != "Oceania") %>% # everything except Oceania
  droplevels() #drop the unused level (i.e., Oceania). 

levels(no_oceania)
```

As shown above, Oceania is not one of the continents in the *no_oceania* dataset.

## 2.2 Reordering levels
Let us plot the distribution of life expectancy of the four continents without Oceania.

```{r 2.1}
no_oceania %>%
  group_by(continent) %>%
  ggplot(aes(continent, lifeExp)) +
  geom_violin()

```

Now let's try to reorder the x-axis. How about having the one with the largest population comes first? Let's do this.

```{r 2.1}
no_oceania %>%
  group_by(continent) %>%
  ggplot() +
  geom_violin(
    aes(fct_reorder(continent, pop, sum, .desc = TRUE), # for each continent, sum(pop) and then sort in the ASCENDING order
        lifeExp)
    )

```

As we all know, Asia has the largest population followed by Americas, Africa and Europe. So the violin plots are presented in that order.

# Exercise 3: File I/O
In this exercise, I'm going to summarize the life expectancy data in 'no_oceania' tibble, which I've just generated above, and  export it to disk. And then I'll reload the exported file.

First, let's summarize the tibble so that the average and standard deviation of life expectancy are calculated by countires. After that, let's export the data.

```{r 3.1}
no_oceania %>%
  group_by(country) %>%
  summarize(averageLifeExp = mean(lifeExp),
            sdLifeExp = sd(lifeExp)) %>%
  write_csv(here::here("hw05", "data","ex03_lifeExp.csv"))

```

Now let's import the exported .csv file from the disk!

```{r 3.2}
imported <- read_csv(here::here("hw05", "data","ex03_lifeExp.csv"))
imported

```

Looking good, except for the fact that the country is not a *factor* variable as shown below! So the datatype didn't survive the round trip of writing to file then reading back in.

```{r 3.3}
class(imported$country)
```

However, when we re-order the country by the averageLifeExp, it would be changed into *factor* before reordered.

```{r 3.4}
imported %>% 
  mutate(country = fct_reorder(country, averageLifeExp, .desc = TRUE)) %>% 
  arrange(country)
```

# Exercise 4: Visualization design

Create a side-by-side plot and juxtapose your first attempt.



# Exercise 5: Writing figures to file

Use ggsave()
