---
title: "Hw01_Ex02 R Markdown for Gapminder exploration"
author: "Stanley Nam"
date: "September 17, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 2
always_allow_html: yes
---

<!-- TODO: more tinkering with the YAML header??? -->

```{r load, warning = FALSE, echo = FALSE}
library(gapminder)
```

# First things first: view the whole dataset

```{r whole data}
tibbleData <- tibble::as_tibble(gapminder)
DT::datatable(tibbleData)
```

# What I want to do

Investigate factors behind the different life expectancies using contemporary data.

# Samples later than 2000

We are living in the 21st century. Why not give more weight on the contemporary data (from 2001)?
```{r contemporary data}
contempData <- gapminder[which(gapminder$year > 2000),]
```

Now I'm a bit concerned: do we have enough data points? Provide descriptive statistics of the dataset.
```{r summary }
summary(contempData)
```

# Factors that influence life expectancy

## Correlation between GDP per capita and life expectancy

...although it seems obvious that the rich live longer.

Plot GDP on the y-axis and Life Expectancy on the x-axis:

```{r gdp-lifeexp, echo=FALSE}
GDP_per_capita <- contempData$gdpPercap
life_Exp <- contempData$lifeExp
plot(GDP_per_capita, life_Exp)
```

## correlation between population and Life Expectancy

This one is less obvious, but research (Calhoun, J.B., 1973. Death squared: the explosive growth and demise of a mouse population. Proc. R. Soc. Med. 66, 80–88.) may indicate that larger population (higher population density, to be honest) has a negative effect on the life expectancy.

Plot population on the y-axis and Life Expectancy on the x-axis:

```{r pop-lifeexp, echo=FALSE}
population <- contempData$pop 
plot(population, life_Exp)
```

