---
title: "Hw01_Ex03 R Markdown for Gapminder exploration"
author: "Stanley Nam"
date: "September 17, 2019"
output: ioslides_presentation
---

```{r load, warning = FALSE, echo = FALSE}
library(gapminder)
```

# Raw data

## First things first: the whole dataset

```{r whole data, echo = FALSE}
tibbleData <- tibble::as_tibble(gapminder)
DT::datatable(tibbleData, extensions = 'Scroller', options = list(
  deferRender = TRUE,
  scrollY = 340,
  scroller = TRUE
))

```

# Objectives

## What I want to do

Investigate factors behind the different life expectancies using contemporary data.

## Samples later than 2000

We are living in the 21st century. Why not give more weight on the contemporary data (from 2001)?
```{r contemporary data}
contempData <- gapminder[which(gapminder$year > 2000),]
head(contempData)
```

## Samples later than 2000

Provide descriptive statistics of the dataset.
```{r summary }
summary(contempData)
```

# Factors that influence life expectancy

## Correlation between GDP per capita and life expectancy

...although it seems obvious that the rich live longer.

Plot GDP on the y-axis and Life Expectancy on the x-axis:
```{r gdp-lifeexp, echo=FALSE}
GDP_per_capita <- contempData$gdpPercap
life_Exp <- contempData$lifeExp
plot(GDP_per_capita, life_Exp)
```

## Correlation between population and life expectancy

Calhoun (1973) indicates that higher population density results in a shorter life expectancy.

Let's plot a graph with population on the y-axis and life expectancy on the x-axis (graph presented on the next slide).

## Correlation between population and life expectancy

Messy..
```{r pop-lifeexp, echo=FALSE}
population <- contempData$pop 
plot(population, life_Exp)
```

# Discussion and Conclusion

## Discussion and Conclusion

It is fun to make presentation slides using R Markdown.